How to obtain the redshift distribution from probabilistic redshift estimates

### 2017-09-07 research/paper/draft.tex
updating text for new tests

\documentclass[iop]{emulateapj}

@@

\newcommand{\myemail}{aimalz@nyu.edu}
\newcommand{\textul}{\underline}
\newcommand{\chippr}{\texttt{CHIPPR} }

@@

\altaffiltext{1}{Center for Cosmology and Particle Physics, Department of 
Physics,
  New York University, 726 Broadway, 9th floor, New York, NY 10003, USA}
\altaffiltext{2}{Simons Center for Computational Astrophysics, 162 Fifth 
Avenue, 7th floor, New York, NY 10010, USA}
\altaffiltext{3}{Center for Data Science, New York University, 60 Fifth Avenue, 
7th floor, New York, NY 10003, USA}
\altaffiltext{4}{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 
Heidelberg, Germany}
\altaffiltext{5}{[SLAC]}

\begin{abstract}
The redshift distribution $n(z)$ is a crucial ingredient for weak lensing 
cosmology.  Spectroscopically confirmed redshifts for the dim and numerous 
galaxies observed by weak lensing surveys are expected to be inaccessible, 
making photometric redshifts (photo-$z$s) the next best alternative.  Because 
of the nontrivial inference involved in their determination, photo-$z$ point 
estimates are being superseded by photo-$z$ probability distribution functions 
(PDFs).  However, analytic methods for utilizing these new data products in 
cosmological inference are still evolving.  This paper presents a novel 
approach to estimating the posterior distribution over $n(z)$ from a survey of 
galaxy photo-$z$ PDFs based upon a probabilistic graphical model of 
hierarchical inference.  We present the Cosmological Hierarchical Inference 
with Probabilistic Photometric Redshifts (\chippr) code implementing this 
technique, as well as its validation on mock data and testing on the 
\texttt{Buzzard} simulations.  \chippr yields a more accurate characterization 
of $n(z)$ containing information beyond the best-fit estimator produced by 
traditional procedures.  The publicly available code is easily extensible to 
other one-point statistics that depend on redshift.

@@

After a brief literature review addressing how photo-$z$ PDFs are currently 
used in cosmology, this paper aims to answer the following questions:

\begin{itemize}
	\item Why should we question existing methods?
	\item How can we improve the effectiveness of using photo-$z$ PDFs in 
inference?
	\item How does the result of \chippr compare to established estimators 
in terms of the accuracy of $n(z)$?
	\item How significant is the effect of the discrepancy between $n(z)$ 
estimators on cosmological constraints?
\end{itemize}

@@

The publicly available \chippr code implements the probabilistic graphical 
model presented in Sec. \ref{sec:model}.

In addition to the choices and assumptions underlying the probabilistic 
graphical model, the implementation of \chippr makes choices and assumptions of 
its own.

\begin{enumerate}
	\item \chippr is only applicable if the redshift interim posteriors and 
interim prior are accurate.
	\item \chippr currently only accepts photo-$z$ PDFs and produces $n(z)$ 
samples of a single format, that of the piecewise constant parametrization, 
also referred to as a binned histogram parametrization and a sum of top hat 
functions.
\end{enumerate}

@@

We demonstrate the superiority of \chippr over alternative approaches in a 
number of compelling test cases on mock data.  Each experiment is characterized 
by a single change to a fiducial case in order to isolate the influence of 
systematic effects known to be relevant to photo-$z$ estimation and propagation 
in analysis.

@@

The mock data in these tests consists of photo-$z$ interim posteriors rather 
than photometric data because the various existing methods for deriving 
photo-$z$ interim posteriors do not in general yield results that are 
consistent with one another, indicating that their systematics are not 
well-understood.  Because the mock data is independent of any choice of 
photo-$z$ PDF production method, we not only ensure that our photo-$z$ interim 
priors are perfectly understood but also deter readers from assuming that 
\chippr has any preference over the method by which photo-$z$ interim 
posteriors are derived from photometric data.

@@

@@

\begin{enumerate}
	\item \label{itm:true} Choose a true $n(z)$ that is a continuous 
function with known parameters $\vec{\phi}'$.
	\item \label{itm:sample} Sample the true $n(z)$ to create a catalog of 
$N$ true redshifts $z'_{i}$.
	\item \label{itm:scatter} Choose a true intrinsic scatter parameter 
$\sigma$, and sample Gaussians $p(z''_{i} | z'_{i}, \sigma) = 
\mathcal{N}(z'_{i}, \sigma)$ to get "observed" redshifts $z'_{i}$ defining 
Gaussian likelihoods $p(\vec{d}_{i} | z''_{i}, \sigma) = \mathcal{N}(z''_{i}, 
\sigma)$.
	\item \label{itm:interim} Choose a parametrization and the parameters 
$\vec{\phi}^{*}$ of the interim prior $p(z_{i} | \vec{\phi}^{*})$.
	\item \label{itm:posterior} Create interim posteriors $\log[p(z_{i} | 
\vec{d}_{i}, \vec{\phi}^{*})] = \log[p(\vec{d}_{i} | z'_{i})]\ \log[p(z_{i} | 
\vec{\phi}^{*})]$ in this parametrization.
\end{enumerate}

@@

The Kullback-Leibler divergence is our primary measure of the accuracy of 
estimators of $n(z)$ in cases of mock data with known true redshifts.

\textbf{Review precision and bias from \texttt{kld.ipynb} and interpret in 
terms of a \% difference.}

@@

Existing $n(z)$ estimation routines will always produce a biased estimator when 
the interim prior is not equal to the true $n(z)$.  We demonstrate here that 
regardless of the appropriateness of the interim prior as an approximation to 
the true $n(z)$, \chippr is not affected by the choice of the interim prior so 
long as it has nontrivial coverage in the space of redshift.  These tests 
modify Step \ref{itm:interim} of Sec. \ref{sec:mockdata}.

@@

To show how the choice of $n(z)$ estimator propagates to cosmological 
constraints, we apply \chippr to a data from a realistic cosmological 
simulation (probably \texttt{Buzzard}) with photo-$z$ PDFs produced by a 
popular method (probably \texttt{BPZ}).

@@

As in Sec. \ref{sec:validintro}, the mock data takes the form of photo-$z$ 
interim posteriors, but that is where the similarity ends.  These photo-$z$ 
interim posteriors are derived from the photometry resulting from the 
\texttt{Buzzard} simulation by way of \texttt{BPZ}.  Because the simulation 
begins with setting true values of the cosmological parameters, we can 
propagate the different estimators of $n(z)$ through a forecasting code (which 
one?) to generate error ellipses on the cosmological parameters.

\subsubsection{Mock Data}
\label{sec:buzzard}

We summarize the details of the \texttt{Buzzard} simulation here.

@@

Because the realistic mock data of Sec. \ref{sec:buzzard} is associated with 
true values of the cosmological parameters, we may compare the quality of 
cosmological constraints under different estimators of $n(z)$, creating a 
figure like Fig. \ref{fig:scgsr}.  We perform this analysis using a forecasting 
code that takes $n(z)$ and produces projected error ellipses in the space of 
cosmological parameters.

@@

\begin{itemize}
	\item Existing $n(z)$ estimation methods produce biased estimators that 
propagate to inaccuracies in characterizing the cosmological parameters.
	\item Photo-$z$ PDFs are probabilistic data products so must be handled 
in a mathematically consistent manner such as the probabilistic graphical model 
outlined in this paper.
	\item In addition to coming with its own error distribution, the $n(z)$ 
estimator produced by \chippr is quantifiably more accurate than established 
estimators.
	\item Propagation of the \chippr result leads to a quantifiable 
improvement in the constraints on cosmological parameters.
\end{itemize}

In conclusion, we discourage the community from continuing to use the stacked 
estimator and reductions of photo-$z$ PDFs to redshift point estimates in 
obtaining estimators of $n(z)$.  \chippr is freely available to the community 
for incorporation into evolving data analysis pipelines.

\appendix{}

>>>>

\documentclass[12pt, onecolumn]{emulateapj}

@@

\newcommand{\myemail}{aimalz@nyu.edu}
\newcommand{\textul}{\underline}
\newcommand{\chippr}{\texttt{chippr}}

@@

\altaffiltext{1}{Center for Cosmology and Particle Physics, Department of 
Physics,
  New York University, 726 Broadway, 9th floor, New York, NY 10003, USA}
\altaffiltext{2}{Simons Center for Computational Astrophysics, 162 Fifth 
Avenue, 7th floor, New York, NY 10010, USA}
\altaffiltext{3}{Center for Data Science, New York University, 60 Fifth Avenue, 
7th floor, New York, NY 10003, USA}
\altaffiltext{4}{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 
Heidelberg, Germany}
\altaffiltext{5}{SLAC National Accelerator Laboratory, Menlo Park, CA 94025, 
USA}

\begin{abstract}
The redshift distribution $n(z)$ is a crucial ingredient for weak lensing 
cosmology.  Spectroscopically confirmed redshifts for the dim and numerous 
galaxies observed by weak lensing surveys are expected to be inaccessible, 
making photometric redshifts (photo-$z$s) the next best alternative.  Because 
of the nontrivial inference involved in their determination, photo-$z$ point 
estimates are being superseded by photo-$z$ probability distribution functions 
(PDFs).  However, analytic methods for utilizing these new data products in 
cosmological inference are still evolving.  This paper presents a novel 
approach to estimating the posterior distribution over $n(z)$ from a survey of 
galaxy photo-$z$ PDFs based upon a probabilistic graphical model of 
hierarchical inference.  We present the Cosmological Hierarchical Inference 
with Probabilistic Photometric Redshifts (\chippr) code implementing this 
technique, as well as its validation on mock data and testing on realistic mock 
data.  \chippr\ yields a more accurate characterization of $n(z)$ containing 
information beyond the best-fit estimator produced by traditional procedures.  
The publicly available code is easily extensible to other one-point statistics 
that depend on redshift.

@@

\textbf{Obviously I need to write the lit. review!}

This paper aims to answer the following questions:

\begin{itemize}
	\item Why should we question existing methods?
	\item How can we improve the effectiveness of using photo-$z$ PDFs in 
inference?
	\item How does the result of \chippr\ compare to established estimators 
in terms of the accuracy of $n(z)$?
	\item How significant is the effect of the discrepancy between $n(z)$ 
estimators on cosmological constraints?
\end{itemize}

@@

The publicly available \chippr\ code implements the probabilistic graphical 
model presented in Sec. \ref{sec:model}.

In addition to the choices and assumptions underlying the probabilistic 
graphical model, the implementation of \chippr\ makes choices and assumptions 
of its own.

\begin{enumerate}
	\item \chippr\ is only applicable if the redshift interim posteriors 
and interim prior are accurate.
	\item \chippr\ currently only accepts photo-$z$ PDFs and produces 
$n(z)$ samples of a single format, that of the piecewise constant 
parametrization, also referred to as a binned histogram parametrization and a 
sum of top hat functions.
\end{enumerate}

@@

We demonstrate the superiority of \chippr\ over alternative approaches in a 
number of compelling test cases on mock data.  Each experiment is characterized 
by a single change to a fiducial case in order to isolate the influence of 
systematic effects known to be relevant to photo-$z$ estimation and propagation 
in analysis.

@@

The mock data in these tests consists of photo-$z$ interim posteriors rather 
than photometric data because the various existing methods for deriving 
photo-$z$ interim posteriors do not in general yield results that are 
consistent with one another, indicating that their systematics are not 
well-understood.  Because the mock data is independent of any choice of 
photo-$z$ PDF production method, we not only ensure that our photo-$z$ interim 
priors are perfectly understood but also deter readers from assuming that 
\chippr\ has any preference over the method by which photo-$z$ interim 
posteriors are derived from photometric data.

@@

\textbf{This needs reorganization and clarification.}

@@

\begin{enumerate}
	\item \label{itm:true} Choose a true $n(z)$ that is a continuous 
function with known parameters $\vec{\phi}'$.
	\item \label{itm:consistent}  Choose an empirical distribution 
$p(z_{spec}, z_{phot} | \textul{\sigma}, \vec{\phi}')$ emulating the error 
model with parameters $\textul{\sigma}$ of photo-$z$s $z_{phot}$ and 
spectroscopic redshifts $z_{spec}$.
	\item \label{itm:sample} Sample $N$ pairs $(z_{phot}, z_{spec})$ from 
$p(z_{spec}, z_{phot} | \textul{\sigma}, \vec{\phi}')$.
	\item \label{itm:interim} Choose an interim $n(z)$ with known 
parameters $\vec{\phi}^{*}$ and construct its empirical distribution 
$p(z_{spec}, z_{phot} | \textul{\sigma}, \vec{\phi}^{*})$.
	\item \label{itm:propagate} Follow Eq. \ref{eq:transformation} to 
create the interim empirical distribution.
	\item \label{itm:posterior} Create interim posteriors $p(z | z_{phot}, 
\textul{\sigma}, \vec{\phi}^{*})$ by evaluating the interim empirical 
distribution at values of $z_{phot}$ drawn in It. \ref{itm:sample}.
\end{enumerate}

With a goal of emulating the systematics of traditional $z_{spec}$ vs. 
$z_{phot}$ scatterplots, we start from probability distributions in that space. 
 This means there are two parameters $z_{spec}$ and $z_{phot}$ drawn from a 
two-dimensional distribution $p(z_{spec}, z_{phot} | \textul{\sigma}, 
\vec{\phi})$, where $\textul{\sigma}$ contains parameters concerning the 
relationship between $z_{spec}$ and $z_{phot}$, and $\vec{\phi}$ contains 
hyperparameters concerning $n(z)$.  We know from basic probability that
\begin{align}
\label{eq:trivial}
p(z_{spec}, z_{phot} | \textul{\sigma}, \vec{\phi}) =& p(z_{phot} | z_{spec}, 
\textul{\sigma}, \vec{\phi})\ p(z_{spec} | \textul{\sigma}, \vec{\phi}).
\end{align}
Because $z_{spec}$ depends only on $\vec{\phi}$ and $z_{phot}$ is independent 
of $\vec{\phi}$, this becomes
\begin{align}
\label{eq:simpler}
p(z_{spec}, z_{phot} | \textul{\sigma}, \vec{\phi}) =& p(z_{phot} | z_{spec}, 
\textul{\sigma})\ p(z_{spec} | \vec{\phi}).
\end{align}
We know how to make reasonable choices for both terms on the righthand side of 
Eq. \ref{eq:simpler} and note that this is equivalent to using $z_{phot}$ as a 
proxy for photometric data and thinking of $z_{spec}$ as the $z$ we seek to 
characterize.

In the $z_{spec}$ vs. $z_{phot}$ scatterplot, $p(z_{phot} | z_{spec}, 
\textul{\sigma}, \vec{\phi})$ would be represented by vertical slices, but we 
are interested in the horizontal slices that represent $p(z_{spec} | z_{phot}, 
\textul{\sigma}, \vec{\phi})$.  Now we must distinguish the true 
hyperparameters $\vec{\phi}'$ we wish to estimate from the interim 
hyperparameters $\vec{\phi}^{*}$.  If the horizontal slices carry information 
about $\vec{\phi}$, we would like to get a catalog of $p(z_{spec} | z_{phot}, 
\textul{\sigma}, \vec{\phi}^{*})$ that's consistent with $p(z_{spec} | 
z_{phot}, \textul{\sigma}, \vec{\phi}')$.

We achieve the goal of propagating that information by dividing out the true 
$n(z)$ and multiplying by the interim $n(z)$ as in
\begin{align}
\label{eq:transformation}
p(z_{spec} | z_{phot}, \textul{\sigma}, \vec{\phi}^{*}) =& p(z_{spec} | 
z_{phot}, \textul{\sigma}, \vec{\phi}')\ \frac{p(z_{spec} | 
\vec{\phi}^{*})}{p(z_{spec} | \vec{\phi}')},
\end{align}
since $p(z_{spec} | z_{phot}, \textul{\sigma}, \vec{\phi})$ is separable and 
the term $p(z_{spec} | z_{phot}, \textul{\sigma})$ changes independently of the 
$p(z_{spec} | \vec{\phi})$ term.  The information about $\vec{\phi}'$ is 
retained by the elements of the catalog: only $(z_{spec}, z_{phot})$ pairs 
drawn from the true distribution will be represented in the catalog.

This is the only fully self-consistent way to ensure that a draw from the 
posterior distribution $p(z_{spec} | z_{phot}, \textul{\sigma}, \vec{\phi})$ 
does in fact follow that distribution.

@@

\textbf{This needs explication.}

The Kullback-Leibler divergence (KLD) is our primary measure of the accuracy of 
estimators of $n(z)$ in cases of mock data with known true redshifts.  The KLD
\begin{align}
\label{eq:kld}
\mathrm{KLD} &= \int_{-\infty}^{\infty}\ n(z)\ 
\log\left[\frac{n(z)}{\hat{n}(z)}\right]\ dz
\end{align}
measures the loss of information when using an estimator of a probability 
distribution instead of the true probability distribution, where the 
probability distribution in question is, in our case, the redshift distribution 
function.

\subsubsection{Cosmological Parameter Covariance}
\label{sec:cosmoparams}

We also perform a test with mock data derived from a self-consistent cosmology 
likelihood analysis using 
\texttt{CosmoLike}\footnote{https://github.com/CosmoLike} 
\citep{krause_cosmolike_2017}.

@@

Existing $n(z)$ estimation routines will always produce a biased estimator when 
the interim prior is not equal to the true $n(z)$.  We demonstrate here that 
regardless of the appropriateness of the interim prior as an approximation to 
the true $n(z)$, \chippr\ is not affected by the choice of the interim prior so 
long as it has nontrivial coverage in the space of redshift.  These tests 
modify Step \ref{itm:interim} of Sec. \ref{sec:mockdata}.

@@

To show how the choice of $n(z)$ estimator propagates to cosmological 
constraints, we apply \chippr\ to a data from a realistic cosmological 
simulation.

@@

\textbf{I need to write this section!}

A cosmological analysis is conducted with \texttt{CosmoLike}.

\subsubsection{Mock Data}
\label{sec:realistic}

\textbf{This test will combine several systematics discussed above.}

We summarize the details of the realistic data here.

@@

Because the realistic mock data of Sec. \ref{sec:realistic} is associated with 
true values of the cosmological parameters, we may compare the quality of 
cosmological constraints under different estimators of $n(z)$, creating a 
figure like Fig. \ref{fig:scgsr}.  We perform this analysis using a forecasting 
code that takes $n(z)$ and produces projected error ellipses in the space of 
cosmological parameters.

@@

\begin{itemize}
	\item Existing $n(z)$ estimation methods produce biased estimators that 
propagate to inaccuracies in characterizing the cosmological parameters.
	\item Photo-$z$ PDFs are probabilistic data products so must be handled 
in a mathematically consistent manner such as the probabilistic graphical model 
outlined in this paper.
	\item In addition to coming with its own error distribution, the $n(z)$ 
estimator produced by \chippr\ is quantifiably more accurate than established 
estimators.
	\item Propagation of the \chippr\ result leads to a quantifiable 
improvement in the constraints on cosmological parameters.
\end{itemize}

In conclusion, we discourage the community from continuing to use the stacked 
estimator and reductions of photo-$z$ PDFs to redshift point estimates in 
obtaining estimators of $n(z)$.  \chippr\ is freely available to the community 
for incorporation into evolving data analysis pipelines.
